Query Optimization for Dynamic Imputation

### 2017-05-12 paper/experiments.tex

\subsubsection{Planning Times}
\begin{table}
\centering
\input{tables/planning_summary_exact_pareto.tex}
\caption{Planning times for varying numbers of joins. \todo{update with new table and values}}
  % Planning times are sub-second for up to 4 joins. For 5+  joins planning times rise. Many practical queries
% can be formulated with a smaller number of joins, but to support more complex queries \ProjectName{} provides an
% optimization that maintains approximate Pareto sets for queries with many joins. This reduces the 29 second
% planning time shown for 7 joins by approximately 45\%, on average.
\label{table:planning-times}
\end{table}

\todo{Update: Planning times are reasonable, with a median planning time of \planningmediantime{} ms. over the benchmark queries, constituting \planningruntimepercent{}\% of total runtime.
In all cases, the optimizer returned a query plan within \planningmaxtime{} ms.}
The choice of $\alpha$ has no impact on planning times, because query selection is done after collecting the Pareto frontier.

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits our planning algorithm by constructing a series of queries with increasing number of joins over the CDC tables.
Planning times for queries involving 1--5 joins---a practical value for real world exploratory queries---are less than 1 second.
\todo{update: For queries with 7 joins, planning time increases substantially to 29 seconds, on average.
  It is possible to reduce planning time by approximating the Pareto set, particularly for queries with a large number of joins.
  With this optimization, the queries with 7 joins have an average planning time of 16 seconds (2 seconds SE).}

%\todobox{Jose: I'm not a fan of this line, I would remove}{Furthermore, at that point, it may make more sense to specify a coherent model over the entire base table instead.}

>>>>

\subsubsection{Planning Times}
\begin{figure}
\centering
\includegraphics[scale=0.4]{figures/planning_times_plot.png}
\caption{Planning times are sub-second for up to 5 joins. For more joins, we implement an approximation to Pareto sets. Users can increase the level of
approximation for more aggressive planning reductions.
}
\label{fig:planning-times}
\end{figure}

\todo{Update: Planning times for our benchmark queries are reasonable, with a median planning time of \planningmediantime{} ms. over the benchmark queries, constituting \planningruntimepercent{}\% of total runtime.
In all cases, the optimizer returned a query plan within \planningmaxtime{} ms.}
The choice of $\alpha$ has no impact on planning times, because query selection is done after collecting the Pareto frontier.

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits of our planning algorithm by constructing a series of queries with increasing number of joins over the CDC tables.
Planning times (~\Cref{fig:planning-times}) for queries involving 1--5 joins---a practical value for a practical subset of real world exploratory queries---are less than 1 second.
We extended \ProjectName{} to support an approximation of Pareto sets to improve planning times for queries with 6 or more joins. Approximate Pareto sets
displayed a reduction in planning time linear with respect to the reduction in size of the final frontier. Users can increase the level of approximation to reduce planning times further.
%\todobox{Jose: I'm not a fan of this line, I would remove}{Furthermore, at that point, it may make more sense to specify a coherent model over the entire base table instead.}
